Title Goes Here

### 2021-02-19 analysis/paper/paper.Rmd
Readme and data download

geom_contour_pval <- function(mapping, p.value = 0.01, size = 0.1) {
  list(
    stat_contour2(mapping, breaks = p.value, geom = pattern_dots),
    geom_contour2(mapping, breaks = p.value, size = size)
  )
}

@@

For the second complex principal component, the phase was chosen so that the coefficent of determination between the Oceanic Niño Index [@bamston1997] and the real part was minimised. It's important to reassure the reader that this procedure does not create a spurious correlation, it only takes whatever relationship that already exist and aligns it with a specific phase. If the relationship is very weak, the rotation cannot generate a strong relationship.

@@

pc2[psa_time, on = "time"] %>% 
  .[PC != "SAM"] %>% 
  copy() %>% 
  .[, hgt := rotate(hgt, psa_angle)] %>%
  sep_ReIm() %>% 
  ggplot(aes(value, hgt)) +
  geom_point() +
  scale_y_continuous("EOF2") +
  scale_x_continuous("PSA") +
  facet_grid(part ~ PC)
```

@@

sam_r2 %>%
  ggplot(aes(lev, estimate^2)) +
  geom_point(data = ~.x[p.adjust(p.value, "fdr") < 0.01], aes(color = term)) +
  geom_line(aes(color = term, size = term)) +
  scale_x_level() +
  scale_y_continuous("r²", limits = c(0, 1)) +
  scale_size_manual(guide = "none",
                    values = c(full = 1.2,
                               asym = 0.5,
                               sym = 0.5)) +
  scale_color_brewer(NULL, palette = "Dark2", labels = c(full = "SAM",
                                                         asym = "A-SAM",
                                                         sym = "S-SAM")) +
  facet_grid(part ~ PC, labeller = labeller(PC = c(PC1 = "EOF1", PC2 = "EOF2"))) +
  coord_flip() +
  tag_facets("rc")
```

@@

rotations_o3 %>%
  ggplot(aes(rotation, cor)) +
  geom_vline(xintercept = best_rotation_pc1) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_line(aes(color = part), size = 1) +
  scale_x_continuous("Rotation", breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous("Correlation with O3 zonal wave 1") +
  scale_color_brewer(NULL, palette = "Dark2", labels = lab_cplx)
```

```{r rotations-pc2, fig.cap = "R2 between EOF2 Real and Imaginary parts and ONI index for different rotation parameters of the real and imaginary parts."}
rotations_pc2 %>%
  melt(id.vars = "rotation") %>%
  ggplot(aes(rotation, value^2)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_vline(xintercept = best_rotation_pc2) +
  geom_line(aes(color = variable), size = 1) +
  scale_x_continuous("Rotation", breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous("r^2") +
  scale_color_brewer(NULL, palette = "Dark2", labels = lab_cplx)
```

>>>>

geom_contour_pval <- function(mapping, p.value = 0.01, size = 0.1) {
  list(
    stat_contour_fill(mapping, breaks = c(0, p.value), fill = NA, geom = pattern_dots),
    geom_contour2(mapping, breaks = p.value, size = size)
  )
}

@@

For the second complex principal component, the phase was chosen so that the coefficient of determination between the Oceanic Niño Index [@bamston1997] and the real part was minimised. It's important to reassure the reader that this procedure does not create a spurious correlation, it only takes whatever relationship that already exist and aligns it with a specific phase. If the relationship is very weak, the rotation cannot generate a strong relationship.

We calculated the PSA1 and PSA2 modes as the third and fourth leading EOF of seasonal mean 500 hPa geopotential height with all seasons together following @mo2001.

@@

pc2[psa_time, on = "time"] %>% 
  .[PC != "SAM"] %>% 
  copy() %>% 
  # .[, hgt := rotate(hgt, psa_angle)] %>%
  sep_ReIm() %>% 
  ggplot(aes(value, hgt)) +
  geom_point() +
  scale_y_continuous("EOF2") +
  scale_x_continuous("PSA") +
  facet_grid(part ~ PC)
```

@@

sam_r2 %>%
  ggplot(aes(lev, estimate^2)) +
  geom_vline(xintercept = c(50, 200),  size = 0.2, color = "gray50") +
  geom_point(data = ~.x[p.adjust(p.value, "fdr") < 0.01], aes(color = term)) +
  geom_line(aes(color = term, size = term)) +
  scale_x_level() +
  scale_y_continuous("r²", limits = c(0, 1)) +
  scale_size_manual(guide = "none",
                    values = c(full = 1.2,
                               asym = 0.5,
                               sym = 0.5)) +
  scale_color_brewer(NULL, palette = "Dark2", labels = c(full = "SAM",
                                                         asym = "A-SAM",
                                                         sym = "S-SAM")) +
  facet_grid(part ~ PC, labeller = labeller(PC = c(PC1 = "EOF1", PC2 = "EOF2"))) +
  coord_flip() +
  tag_facets("rc")
```

@@

rotations_o3 %>%
  ggplot(aes(rotation, cor)) +
  geom_vline(xintercept = best_rotation_pc1) +
  geom_hline(yintercept = 0,  size = 0.2, color = "gray50") +
  geom_line(aes(color = part), size = 1) +
  scale_x_continuous("Rotation", breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous("Correlation with O3 zonal wave 1") +
  scale_color_brewer(NULL, palette = "Dark2", labels = lab_cplx)
```

```{r rotations-pc2, fig.cap = "R2 between EOF2 Real and Imaginary parts and ONI index for different rotation parameters of the real and imaginary parts."}
rotations_pc2 %>%
  melt(id.vars = "rotation") %>%
  ggplot(aes(rotation, value^2)) +
  geom_hline(yintercept = 0,  size = 0.2, color = "gray50") +
  geom_vline(xintercept = best_rotation_pc2) +
  geom_line(aes(color = variable), size = 1) +
  scale_x_continuous("Rotation", breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous("r^2") +
  scale_color_brewer(NULL, palette = "Dark2", labels = lab_cplx)
```
